---
title: "Thurs_pm_fusion"
author: "Megan Cattau"
date: "June 23, 2016"
output: html_document
---

Load libraries
```{r load_library}

# Load the libraries
library(raster)
library(rhdf5)
library(rgdal)
library(neonAOP)
```

Import LiDAR data
```{r import_lidar_data}

# import digital surface model (DSM)
dsm<-raster("../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarDSM.tif")

# import digital terrain model (DTM)
dtm<-raster("../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarDTM.tif")

# import canopy height model (chm)
chm<-raster("../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif")
plot(chm, main="Canopy Height\n Lower Teakettle, CA")

# look at a histogram of the CHM
hist(chm, main="Distribution on Canopy Height\n Teakettle, CA", xlab="Tree height (m)", col="springgreen")

# look at loverall image stats. You can't just use mean() because R will then think you want a raster
cellStats(chm, max) # max tree height is 55.7
cellStats(chm, mean) # max tree height is 5.6
cellStats(chm, sd) # max tree height is 10.1

# set 0 values to no data
# chm[chm==0]<-NA
```

Stack rasters together
```{r stack_rasters}
# stack rasters together
lidar.brick<-brick(dsm, dtm, chm)

# Look at it
plot(lidar.brick)

# you can write a function to pass a vector of data
# use calc command to apply across entire scene
# e.g., calculating trend lines
```

Import Hyperspectral data
```{r import_HI_data}

f<-"../NEONdata/D17-California/TEAK/2013/spectrometer/reflectance/Subset3NIS1_20130614_100459_atmcor.h5"

# id the projection
# define the CRS definition by EPSG (these correcpond to UTM zones)
epsg<-32611
```

Compute NDVI from HI 
```{r HA_NDVI}

# pull out some of the bands
bands<-c(60,83)

# look at wavelengths
wavelengths<-h5read(f, "wavelength") 

# read in bands
ndvi.stack<-create_stack(f, 
                         bands=bands,
                         epsg = epsg)

# plot_to_check
plot(ndvi.stack, main="Bands 60 and 83\n Teakettle, CA")

# Compute NDVI
ndvi<-(ndvi.stack[[2]]-ndvi.stack[[1]])/(ndvi.stack[[2]]+ndvi.stack[[1]])

# Rename to something informative
names(ndvi)<-"TEAK_hsiNDVI"

# Plot this to check
plot(ndvi, main="NDVI \n Lower Teakettle, CA")
```

Combine LiDAR and HI NDVI
```{r combine_LiDAR_and_HI}

# try stacking and see what happens - different extent
all.data<-brick(ndvi, lidar.brick)

# Check the extents
extent(ndvi)
extent(lidar.brick)

# Check the extents and crop to smaller raster
if(extent(lidar.brick)==extent(ndvi)) {
} else {
    print("Extents are different, cropping data")
    overlap<-intersect(extent(lidar.brick), extent(ndvi))
    # now let's crop the lidar data to the HSI
    lidar.brick<-crop(lidar.brick, overlap)
    ndvi<-crop(ndvi, overlap)
  }

# 'intersect' can be from different packages, so be explicit about what packages to use

# now stack
all.data<-brick(ndvi, lidar.brick)

#rename things
names(all.data)<-c("NDVI", "DSM", "DTM", "CHM")
```


Compare HI-derived NDVI with NEON NDVI product

```{r compare_HI_NEON_NDVIs}

# Import NEON NDVI product
ndvi.neon<-raster("../NEONdata/D17-California/TEAK/2013/spectrometer/veg_index/TEAK_NDVI.tif")

# do extents match?
extent(ndvi.neon)==extent(all.data)

# Differencing
ndvi.diff<- ndvi-ndvi.neon

# Visulaize
plot(ndvi.diff, main="NDVI Difference \n Teakettle, CA")

# different bc NEON product is average across multiple bands to be moe comparable to Landsat or something

```

Function to compare extents and crop layers if they are different
```{r same_extent}

# write a function to compare extents and crop layers if they are different
same_extent<-function(raster1, raster2) {
  if (extent(raster1)==extent(raster2)) {
    print("Rasters have same extent")
  } else {
    overlap<-raster::intersect(extent(raster1), extent(raster2))
    # crop both rasters
    # might be good to check which is bigger and compare
    print("Extents are different, Cropping data")
    raster1<-crop(raster1, overlap)
    raster2<-crop(raster2, overlap)
    # create a stack of the new rasters
  }
    raster.stack<-stack(raster1, raster2)
    return(raster.stack)
}


# check our data
same_extent(ndvi, lidar.brick)
```

Do we see taller, greener trees on North-facing slopes?
```{r make_NS_object}

# Import NEON aspect (angle a slope is facing)
aspect<-raster("../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarAspect.tif")

# crop aspect to other data
all.data<-same_extent(aspect, all.data)

# rename
names(all.data)<-c("Aspect", "NDVI", "DSM", "DTM", "CHM")

# terrain function in raster package that can calculate aspect for you

# plot aspect
names(all.data)
plot(all.data$Aspect)
plot(all.data$Aspect>270)

# create a classified aspect intermediate output
# first create a matrix of values that represents the classification ranges
# north facing = 1
# south facing = 2

class.m<-c(-1,45,1,
           45,135,NA,
           135,225,2,
           225,315,NA,
           315,360,1)
# bottom edge is non-inclusive, so if you start at 0, 0 doesn't get reclassified

# turn into matrix
rcl.m<-matrix(class.m, ncol=3, byrow=TRUE)

# classify aspect using classification matrix
asp.ns<-reclassify(all.data$Aspect, rcl.m)

plot(asp.ns)
summary(asp.ns)

# get map extent
ns.extent<-extent(asp.ns)

plot(asp.ns,
     col=c("blue", "green"),
     axes=FALSE,
     main="North and South facing slopes \n Teakettle, 
     CA", bty="n",
     legend=FALSE
     )
par(xpd=TRUE) # let it write outside boundaries
legend((par()$usr[2]+20), ns.extent@ymax-100, 
        legend=c("North", "South"),
       fill=c("blue", "green"),
       bty="n")
par(xpd=FALSE) #set it back
```

Create N- and S- facing object and mask data w it
```{r N facing object}

# create N-facing object
north.facing<-asp.ns== 1
north.facing[north.facing==0]<-NA
plot(north.facing)

# create S-facing object
south.facing<-asp.ns== 2
south.facing[south.facing==0]<-NA
plot(south.facing)


```

# histogram of tree height
hist(all.data$CHM, main="Distribution of Canopy Height Model (CHM) values \n Teakettle, CA", col="cornflowerblue")

# get stats of all layers
all.data.stats<- data.frame(t(summary(values(all.data), na.rm=TRUE)))
# t transposes
str(all.data)

ht.mean<-cellStats(all.data$CHM, mean, na.rm=T)
ht.sd<-cellStats(all.data$CHM, sd, na.rm=T)

# @ is for objects (e.g., can't use @ on a dataframe) and $ is for names element inside

# "Tall" objects are going to be mean + 1 sd
# Make a thresholds object
thresholds<-data.frame(id=1)
thresholds$height<-ht.mean+ht.sd

# time for NDVI
hist(all.data$NDVI, col="cornflowerblue", main="Distribution of NDVI values \n Teakettle")

# What counts as "green"
greenRange<-cellStats(all.data$NDVI, range)
greenRange<- greenRange[2]-greenRange[1]

thresholds$green<-cellStats(all.data$NDVI, max)-greenRange/3

# How many N- and S-facing pixels are there?
north.count<-freq(asp.ns, value=1)
south.count<-freq(asp.ns, value=2)

# How many are tall and green?
north.tall.green<-asp.ns==1 & all.data$NDVI>=thresholds$green & all.data$CHM>=thresholds$height

plot(north.tall.green)

# assign zeros to NAs
north.tall.green[north.tall.green==0]<-NA

# how many pixels is that?
freq(north.tall.green, value=1) #5508


# same for S-facing
# How many are tall and green?
south.tall.green<-asp.ns==2 & all.data$NDVI>=thresholds$green & all.data$CHM>=thresholds$height

plot(south.tall.green)

# assign zeros to NAs
south.tall.green[south.tall.green==0]<-NA

# how many pixels is that?
freq(south.tall.green, value=1) #2400
